---
title: "Investing 101: A visual and predictive guide for the rookie investor - Wee Kien"
description: |
  Existing financial data websites such as Yahoo Finance do a good job in providing historical price data and technical indicators, but the beginner investor lacks the knowledge to properly utilise and benefit from these. In addition, we have also identified several gaps in such websites.
  
  For one, these websites do not provide tools to allow the user to compare stocks meaningfully or zoom in to the statistical properties of financial returns. For example, a user is unable to conduct correlation analysis or visualize the distribution of returns. Secondly, these websites also do not provide any form of forecasting to aid in investors’ decisions.
    
  This project aims to improve on the current offering of financial data websites through, Exploratory Data Analysis, Time-Series Forecasting and also Time-Series Clustering.
  
author:
  - name: Wee Kien, Ng (Vincent)
    url: https://www.linkedin.com/in/ngweekiensg/
date: 03-24-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Overview

### 1.1 Background

Traditional financial data websites usually shy away from giving recommendations or advising users on investing decisions. While in depth studies are often conducted and used in academia or research, the outputs of these studies are often not accessible nor timely enough for the regular user. This means that these websites and their applications are limited in their use-cases. This can be due to issues arising from financial or legal liability if users act on the recommendations coming from these financial websites. As such we feel that this a gap between what is available publicly and user requirements. 

### 1.2 Objective and Motivation

TO BE ADDED

To talk about what other modules will focus on.

What we are doing in this particular module.

- Charting and technical analysis
- Correlation analysis between selected portfolio of stocks vs a benchmark index
- Portfolio analysis:
  - Individual portfolio returns
  - Equity curve/ portfolio growth over time


### 1.3 Charting and Technical Analysis


### 1.4 Correlation Analysis


### 1.5 Portfolio Analysis






## 2. Proposed Visualization

### 2.1 Sketch of Proposed Visualization

TO INSERT SKETCH HERE, can be multiple pictures.
Can take from powerpoint slide.


### 2.2 Suggested Interactivity in the Proposed Visualization

TO DESCRIBE in words, the various portions of the interactivity.


## 3. Building the Visualization

### 3.1 Setting up the environment/ packages

First, we run this first line of code to clear the environment and remove existing R objects (if any)

```{r, echo=TRUE, eval=TRUE, fig.cap="Starting on a clean slate"}

rm(list=ls())

```

This code chunk checks if required packages are installed. If they are not installed, the next line of code will install them. The following line is then use to import the library into the current working environment.

```{r, echo=TRUE, eval=TRUE, fig.cap="Installing and importing libraries"}

packages = c('tidyverse','tidyquant','dygraphs','plotly','ggplot2')
for (p in packages) {
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}

```

### 3.2 Charting and Technical Analysis



### 3.3 Correlation Analysis



### 3.4 Portfolio Analysis







## 4. Conclusions and Insights


## 5. References

















